---
title: "Assignment 5"
output: pdf_document
latex_engine: xelatex
header-includes:
- \usepackage{blkarray}
- \usepackage{amsmath}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(ggplot2)
library(rsample)
library(smotefamily)
library(skimr)
library(GGally)
library(dplyr)
library(caret)
library(MASS)
library(car)
library(glmnet)
```

Logistic regression: Predicting a Legendary Pokemon

```{r}
rm(list=ls())
poke_data <- read.csv("pokemon_merged.csv")
head(poke_data)
```

The merged dataset contains duplicate columns, drop duplicate columns
```{r}
num_cols = c('Generation', 'HP','Attack','Defense','Sp..Atk', 
             'Sp..Def','height','weight','Speed','stamina')
cat_cols = c('Type.1','Type.2','pokemon_family','wild_avail',
             'egg_avail','category','research_avail','region')
bool_cols = c('Legendary', 'shiny','shadow')

df_num_cols <- poke_data %>% dplyr::select(all_of(num_cols))
df_cat_cols <- poke_data %>% dplyr::select(all_of(cat_cols))
df_bool_cols <- poke_data %>% dplyr::select(all_of(bool_cols))
poke_clean <- cbind(Name = poke_data$Name, df_num_cols, df_cat_cols, df_bool_cols)

poke_clean[,bool_cols[2:3]] = lapply(poke_clean[,bool_cols[2:3]], 
                                     function(x) ifelse(x == 'No', FALSE, TRUE))
poke_clean[,bool_cols] = poke_clean[,bool_cols] %>% mutate_all(as.logical)
poke_clean[,cat_cols] = poke_clean[,cat_cols] %>% mutate_all(as.factor)

poke_clean$height = as.numeric(gsub(" (.*)$", "", poke_clean$height))
poke_clean$weight = as.numeric(gsub(" (.*)$", "", poke_clean$weight))

poke_clean$shiny <- as.factor(poke_clean$shiny)
poke_clean$shadow <- as.factor(poke_clean$shadow)
poke_clean$Legendary <- as.numeric(poke_clean$Legendary)
head(poke_clean)
```

```{r}
bin_counts <- poke_clean %>%
  count(Legendary) %>%
  mutate(percentage = n / sum(n) * 100)
bin_counts 


#legendary_plot <- ggplot(data = poke_clean, aes(x = Legendary)) +
#  geom_histogram(binwidth = 1, color = "black", fill = "dodgerblue", 
#                 alpha = 0.8, aes(y = ..count.. / sum(..count..) * 100)) +
#  geom_text(data = bin_counts, aes(label = paste0(round(percentage, 1), "%"), 
#                                   y = percentage, x = Legendary), vjust = -0.5) +
#  labs(title = "Histogram of Legendary Pokémon", x = "Legendary", y = "Percentage") +
#  theme_minimal() 
#
#legendary_plot
```
imbalance dataset, and baseline accuracy is 94.7%

logit plot
```{r}
# Create a logistic plot for each variable
colnames(poke_clean)
variables <- c("Attack", "Defense", "Sp..Atk", "Sp..Def", "Speed","Generation","HP","height","weight")

for (var in variables) {
  plot <- ggplot(poke_clean, aes_string(x = var, y = "Legendary")) +
    geom_point() +
    stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE) +
    labs(title = paste("Logistic Regression Plot of", var, "vs Legendary"),
         x = var,
         y = "Legendary (Probability)") +
    theme_minimal()

  print(plot)
}

```

```{r}
variables <- c("Attack", "Defense", "Sp..Atk", "Sp..Def", "Speed", "Generation", "HP", "height", "weight", "stamina")

for (var1 in variables) {
  for (var2 in variables) {
    if (var1 != var2) {
      plot <- ggplot(poke_clean, aes_string(x = var1, y = "Legendary", color = var2)) +
        geom_point() +
        stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE, aes_string(color = var2)) +
        scale_color_gradient(low = "blue", high = "red") +
        labs(title = paste("Logistic Regression Plot of", var1, "and", var2, "vs Legendary"),
             x = var1,
             y = "Legendary (Probability)",
             color = var2) +
        theme_minimal()

      print(plot)
    }
  }
}


```

```{r}
selected_vars <- c("Attack", "Defense", "Sp..Atk", "Sp..Def", "Speed","Generation","HP","height","weight", "Legendary")
train_set_selected <- poke_clean[, selected_vars]

# Convert the 'Legendary' variable to a factor
train_set_selected$Legendary <- as.factor(train_set_selected$Legendary)

plot <- ggpairs(train_set_selected,
                columns = 1:9,
                mapping = aes(color = Legendary),
                lower = list(continuous = wrap("points", alpha = 0.5)),
                diag = list(continuous = wrap("barDiag", alpha = 0.5)),
                upper = list(continuous = wrap("cor", size = 3))) +
  theme_minimal()

print(plot)

```



Based on the EDA part, choose the most related variable to Legendary
check multicollinearity
```{r}
poke_linear <- dplyr::select(poke_clean,Generation, HP, Attack, Defense, Sp..Atk, Sp..Def, height, weight, Speed, stamina, Legendary, Type.1,Type.2,shiny,shadow)
linear_model <- lm(Legendary ~ ., data = poke_linear)
vif_values <- vif(linear_model)
vif_values
```
From Generalized Variance Inflation Factor (GVIF), HP and stamina have high multicollinearity.

```{r}
poke_linear <- dplyr::select(poke_clean,Generation, HP, Attack, Defense, Sp..Atk, Sp..Def, height, weight, Speed, stamina, Legendary, Type.1,Type.2,shiny,shadow)
poke_clean <- dplyr::select(poke_linear, - stamina)
```

```{r}
#class_0 <- sum(poke_clean$Legendary == 1)
#class_1 <- sum(poke_clean$Legendary == 0)
#TN <- 0
#FP <- class_0
#cat("Class 0:", class_0, "\n")
#FN <- 0
#TP <- class_1
#cat("Class 1:", class_1, "\n")
#p <- TP / (TP + FP)
#r <- TP / (TP + FN)
#baseline_f1 <- (2 * p * r) / (p + r)
#
#baseline_f1
#
```




```{r}
set.seed(39)
folds <- createFolds(poke_clean$Legendary, k = 5, list = TRUE, returnTrain = FALSE)

perform_glm <- function(train_data, test_data) {
  glm_model <- glm(Legendary ~ ., data = train_data, family = binomial(link = 'logit'), control = list(maxit = 50))
  preds <- predict(glm_model, newdata = test_data, type = "response")
  pred_class <- ifelse(preds >= 0.5, 1, 0)
  accuracy <- mean(pred_class == test_data$Legendary)
  aic <- AIC(glm_model)
  cm <- confusionMatrix(as.factor(pred_class), as.factor(test_data$Legendary))
  
  return(list(model = glm_model, accuracy = accuracy, aic = aic, cm = cm))
}

cv_results <- lapply(folds, function(test_idx) {
  train_data <- poke_clean[-test_idx, ]
  test_data <- poke_clean[test_idx, ]
  perform_glm(train_data, test_data)
})

best_model_idx <- which.max(sapply(cv_results, function(x) x$accuracy))
best_model <- cv_results[[best_model_idx]]$model
best_accuracy <- cv_results[[best_model_idx]]$accuracy

cat("The best model is:\n")
print(summary(best_model))

cat("\nThe accuracy of the best model is:", best_accuracy)

```


```{r}
set.seed(39)
folds <- createFolds(poke_clean$Legendary, k = 5, list = TRUE, returnTrain = FALSE)

perform_glm <- function(train_data, test_data) {
  glm_model <- glm(Legendary ~ ., data = train_data, family = binomial(link = 'logit'), control = list(maxit = 50))
  preds <- predict(glm_model, newdata = test_data, type = "response")
  pred_class <- ifelse(preds >= 0.5, 1, 0)
  accuracy <- mean(pred_class == test_data$Legendary)
  aic <- AIC(glm_model)
  cm <- confusionMatrix(as.factor(pred_class), as.factor(test_data$Legendary))
  
  return(list(accuracy = accuracy, aic = aic, cm = cm))
}



cv_results <- lapply(folds, function(test_idx) {
  train_data <- poke_clean[-test_idx, ]
  test_data <- poke_clean[test_idx, ]
  perform_glm(train_data, test_data)
})

best_model_idx <- which.max(sapply(cv_results, function(x) x$accuracy))
best_model <- cv_results[[best_model_idx]]

cat("The best model is:\n")
print(best_model)
cat("\nThe accuracy is:", best_model$accuracy)
cat("\nConfusion matrix for the best model:\n")
print(best_model$cm$table)

# Printing the misclassified data
train_data <- poke_clean[-folds[[best_model_idx]], ]
test_data <- poke_clean[folds[[best_model_idx]], ]
glm_model <- glm(Legendary ~ ., data = train_data, family = binomial(link = 'logit'), control = list(maxit = 50))
preds <- predict(glm_model, newdata = test_data, type = "response")
pred_class <- ifelse(preds >= 0.5, 1, 0)

misclassified <- test_data[pred_class != test_data$Legendary, ]
misclassified_as_legendary <- misclassified[misclassified$Legendary == 0, ]
misclassified_as_not_legendary <- misclassified[misclassified$Legendary == 1, ]

cat("\nData misclassified as 'Legendary':\n")
print(misclassified_as_legendary)

cat("\nData misclassified as 'Not Legendary':\n")
print(misclassified_as_not_legendary)
```

```{r}
library(ggplot2)

# Convert the confusion matrix to a data frame
confusion_matrix_df <- as.data.frame(best_model$cm$table)

# Add column names for ggplot
colnames(confusion_matrix_df) <- c("Reference", "Prediction", "Frequency")

# Plot the heatmap of the confusion matrix
ggplot(data = confusion_matrix_df, aes(x = Prediction, y = Reference, fill = Frequency)) +
  geom_tile() +
  geom_text(aes(label = Frequency), color = "white", size = 5) +
  scale_fill_gradient(low = "lightblue", high = "darkblue") +
  labs(title = "Confusion Matrix Heatmap",
       x = "Reference",
       y = "Prediction",
       fill = "Frequency") +
  theme_minimal()

```


```{r}
print(poke_data[139,])
print(poke_data[149,])
print(poke_data[390,])
print(poke_data[631,])
```











